---
title: "Logistic Regression - Imputed Antibiotic Prediciton - Stanford"
author: "Manoj Maddali, MD"
date: "`r Sys.Date()`"
output: html_document
---
  
```{r setup, include=FALSE}
library(tidyverse)
library(bigrquery)
library(gt)
library(caret)
library(xgboost)
library(pROC)
library(MLmetrics)
library(comorbidity)
library(lightgbm)
library(parallel)
library(shapviz)
library(mice)

bq_auth()
```

# Define helper functions
``` {r}
#########################################################################################################################
# Define features for use in model
#########################################################################################################################
categorical_features = c("female", "dialysis", "year", "vasopressor")
numeric_features = c("age", "cci",
                     "wbch", "hgbl", "pltl",
                     "sodiuml", "bicarbl", "bunh", "creath", "glucl",
                     "asth", "alth", "bilih",
                     "albuml",  "lactateh",
                     "hrateh", "rrateh", "temph", "sysbpl", "diasbpl", "weightl",
                     "count_rx_oral", "count_rx_iv", "number_of_admissions",
                     "count_ciprofloxacin_resistance_within_6mo", "count_ciprofloxacin_resistance_before_6mo",
                     "count_oxacillin_resistance_within_6mo", "count_oxacillin_resistance_before_6mo",
                     "count_ampicillin_resistance_within_6mo", "count_ampicillin_resistance_before_6mo",
                     "count_cefazolin_resistance_within_6mo", "count_cefazolin_resistance_before_6mo",
                     "count_ceftriaxone_resistance_within_6mo", "count_ceftriaxone_resistance_before_6mo",
                     "count_cefepime_resistance_within_6mo", "count_cefepime_resistance_before_6mo",
                     "count_pip_tazo_resistance_within_6mo", "count_pip_tazo_resistance_before_6mo",
                     "count_meropenem_resistance_within_6mo", "count_meropenem_resistance_before_6mo",
                     "count_trim_sulfa_resistance_within_6mo", "count_trim_sulfa_resistance_before_6mo",
                     "count_vancomycin_resistance_within_6mo", "count_vancomycin_resistance_before_6mo",
                     "prior_year_antibiogram_ciprofloxacin",
                     "prior_year_antibiogram_oxacillin",
                     "prior_year_antibiogram_ampicillin",
                     "prior_year_antibiogram_cefazolin",
                     "prior_year_antibiogram_ceftriaxone",
                     "prior_year_antibiogram_cefepime",
                     "prior_year_antibiogram_pip_tazo",
                     "prior_year_antibiogram_meropenem",
                     "prior_year_antibiogram_trim_sulfa",
                     "prior_year_antibiogram_vancomycin",
                     "count_prior_pseudomonas_culture",
                     "count_prior_staph_aureus_culture",
                     "count_prior_enterococcus_culture")

#########################################################################################################################
# Calculate and add Charlson comorbidity index scores from parsed ICD9 codes
# Using "comorbidity" package in R
#########################################################################################################################
add_charlson_comorbidity_index = function(df, index) {
  # First parse/explode values of icd9s column for each patient
  icds_parsed = df %>% separate_longer_delim(c(icd9s), delim=",")
  
  # Calculate individual component comorbidity scores
  comorbidity_scores = comorbidity(x=icds_parsed, id=index, code="icd9s", map="charlson_icd9_quan", assign0=TRUE, tidy.codes=TRUE)

  # Score the comorbidity index (unadjusted for age)
  comorbidity_scores$cci = score(comorbidity_scores, assign0 = TRUE)  
  
  # Add the comorbidity index to the original dataframe
  df = df %>% left_join(comorbidity_scores %>% select(all_of(c(index, "cci"))), by=index)
    
  # Adjust the charlson comorbidity index for age
  df = df %>% mutate(cci = cci + ifelse(age >= 80, 4,  
                                        ifelse(age >= 70, 3, 
                                               ifelse(age >= 60, 2, 
                                                      ifelse(age >= 50, 1, 0)))))  
  
  return(df)
}

#########################################################################################################################
# Model training for LightGBM models
#########################################################################################################################
train_test_lgb = function(df, model_antibiotic, features) {
  # Select data for the given antibiotic and features
  df = df %>% filter(antibiotic == model_antibiotic) %>%
    select(csn, resistance, all_of(features)) %>%
    unique()
  
  # Split data into training (pre-2022) and testing (2022 and 2023)
  df_train = df %>% filter(!year %in% c("2022", "2023")) %>% select(-year)
  df_test = df %>% filter(year %in% c("2022", "2023")) %>% select(-year)

  # Set up LightGBM datasets
  df_lgb_train = lgb.Dataset(as.matrix(df_train %>% select(-csn, -resistance)), label = df_train$resistance)
  df_lgb_validation = lgb.Dataset(as.matrix(df_test %>% select(-csn, -resistance)), label = df_test$resistance)

  # Set up hyperparameter grid search
  lgb_grid = expand.grid(
    nrounds = 1000, # early_stopping_rounds will stop training before this if the validation metric doesn't improve
    learning_rate = c(0.01, 0.025, 0.05, 0.1),
    bagging_fraction = c(0.6, 0.8, 0.9, 1),
    bagging_freq = c(5, 10, 20),
    early_stopping_rounds = NA)

  # Set up vector to save each hyperparameter's cross-validated AUC
  cv.aucs = rep(NA, nrow(lgb_grid))

  # Iterate through hyperparameters to find the best cross-validated AUC
  for (i in 1:nrow(lgb_grid)) {
    # Train and test the LightGBM model with current hyperparameters
    model.lgb = lgb.cv(data = df_lgb_train,
                          nrounds = 1000,
                          nfold = 5,
                          params = list(objective = "binary",
                                        metric = "auc",
                                        is_unbalance = TRUE, # Automatically calculate positive weights for unbalanced classes
                                        num_threads = detectCores(), # Use all available cores
                                        learning_rate = lgb_grid[i, "learning_rate"],
                                        max_depth = -1,
                                        bagging_fraction = lgb_grid[i, "bagging_fraction"],
                                        bagging_freq = lgb_grid[i, "bagging_freq"],                                    
                                        early_stopping_rounds = 50,
                                        verbosity = -1))
  
    # Save the AUC
    cv.aucs[i] = model.lgb$best_score
  
    # Save the early_stopping_rounds
    lgb_grid[i, "early_stopping_rounds"] = model.lgb$best_iter
  }

  # Train the final model with the best hyperparameters
  model.lgb = lgb.train(data = df_lgb_train,
                        nrounds = lgb_grid[which.max(cv.aucs), "early_stopping_rounds"],
                        valids = list(validation = df_lgb_validation),
                        params = list(objective = "binary",
                                      metric = "auc",
                                      is_unbalance = TRUE, # Automatically calculate positive weights for unbalanced classes
                                      num_threads = detectCores(), # Use all available cores
                                      learning_rate = lgb_grid[which.max(cv.aucs), "learning_rate"],
                                      max_depth = -1,
                                      bagging_fraction = lgb_grid[which.max(cv.aucs), "bagging_fraction"],
                                      bagging_freq = lgb_grid[which.max(cv.aucs), "bagging_freq"],                                    
                                      verbosity = -1))
  
  # Make predictions
  predictions = predict(model.lgb, as.matrix(df_test %>% select(-csn, -resistance)))
  
  # Calculate ROC
  roc = roc(predictor = predictions, df_test$resistance) 
  
  # Generate confusion matrix
  cm = confusionMatrix(factor(ifelse(predictions >= 0.5, 1, 0)),
                     factor(df_test$resistance), positive = "1", mode = "everything")
  
  # Return the model, AUC, ROC, and confusion matrix
  return(list(antibiotic = str_to_title(model_antibiotic),
              model = model.lgb,
              best_tune = lgb_grid[which.max(cv.aucs),],
              df_train = df_train,
              df_test = df_test,
              predictions = predictions,
              auc = roc$auc,
              roc = roc,
              cm = cm))
}

#########################################################################################################################
# Calculate Shapeley values and plot feature importance
#########################################################################################################################
plot_shap_importance = function(model) {
  shap = shapviz(model$model,
                 X=as.matrix(model$df_train %>% select(-csn, -resistance)),
                 X_pred=as.matrix(model$df_train %>% select(-csn, -resistance)))
  
  shap %>%
    sv_importance(kind="beeswarm") + ggtitle(paste0("SHAP Feature Analysis for ", model$antibiotic, " Model"))
}

#########################################################################################################################
# Plot antibiotic orders saved in next 24 hours
#########################################################################################################################
antibiotics_24h_saved = function(model, df_all_antibiotics_24h, antibiotic24h) {
  # Load test data and predictions from model
  df_test = model$df_test
  df_test$predictions = model$predictions
  
  # Merge with 24h antibiotic order df with df_test
  df_all_antibiotics_24h = df_test %>%
    left_join(df_all_antibiotics_24h %>% select(csn, antibiotic_24h) %>% unique(), by="csn")

  # Filter out the 24h antibiotic of interest
  if (length(antibiotic24h) > 2) {
    stop("Antibiotic24h must only contain max 2 antibiotics")
  }
  
  # Filter out the 24h antibiotic of interest
  if (length(antibiotic24h) == 1) {
    specific_antibiotics_24h = df_all_antibiotics_24h %>%
      filter(antibiotic_24h == str_to_upper(antibiotic24h[1])) %>%
      select(-antibiotic_24h) %>%
      unique()
  } else { # If there are two antibiotics, filter out the common CSNs and take the set of CSNs with both antibiotics
    csn1 = df_all_antibiotics_24h %>% filter(antibiotic_24h == str_to_upper(antibiotic24h[1])) %>% pull(csn)
    csn2 = df_all_antibiotics_24h %>% filter(antibiotic_24h == str_to_upper(antibiotic24h[2])) %>% pull(csn)    
    csns = intersect(csn1, csn2)
    
    specific_antibiotics_24h = df_all_antibiotics_24h %>%
      filter(csn %in% csns) %>%
      select(-antibiotic_24h) %>%
      unique()
  }
  
  # Create a sequence of thresholds to test NPV
  seq = seq(0.00, 1, 0.01)
  excess = data.frame(matrix(ncol = 6, nrow = length(seq)))
  colnames(excess) = c("threshold", "npv", "excess", "relative_excess", "true_excess", "perc_saved")
  excess$threshold = seq
  
  # True excess is the number of orders that were not resistant on culture (gold standard)
  excess$true_excess = nrow(specific_antibiotics_24h %>% filter(resistance == 0))

  # Iterate through thresholds and calculate NPV and excess antibiotic prescriptions
  for (i in seq) {
    # Predict resistance based on threshold
    specific_antibiotics_24h = specific_antibiotics_24h %>% mutate(predicted_resistance = ifelse(predictions >= i, 1, 0))
    # Calculate confusion matrix for NPV
    cm = confusionMatrix(factor(specific_antibiotics_24h$predicted_resistance),
                         factor(specific_antibiotics_24h$resistance),
                         positive = "1", mode = "everything")
    excess[excess$threshold == i, "npv"] = cm$byClass['Neg Pred Value']
    # Calculate number of excess antibiotic orders (predicted resistance = 0, true resistance = 0)
    excess[excess$threshold == i, "excess"] = nrow(specific_antibiotics_24h %>% filter(predicted_resistance == 0, resistance == 0))
    # Calculate relative excess (true excess - excess)
    excess[excess$threshold == i, "relative_excess"] = excess[excess$threshold == i, "true_excess"] - excess[excess$threshold == i, "excess"]
    # Calculate percentage of excess antibiotic orders saved
    excess[excess$threshold == i, "perc_saved"] = round(100*excess[excess$threshold == i, "excess"] / excess[excess$threshold == i, "true_excess"], 2)
  }
  
  # Plot excess antibiotic orders saved vs NPV
  excess %>%
    group_by(npv_group = cut(npv, breaks = seq(0, 1, 0.025))) %>%
    summarize(mean_npv = mean(npv), mean_perc_saved = mean(perc_saved)) %>%
    filter(!is.na(mean_npv)) %>%
    filter(mean_npv > 0.7) %>% 
    ggplot() +
    geom_line(aes(x=mean_npv, y=mean_perc_saved)) +
    geom_vline(xintercept=0.8, linetype="dotted") + 
    geom_ribbon(aes(x=mean_npv, ymin=0, ymax=mean_perc_saved), fill="red", alpha=0.2) +
    xlab("Negative predictive value") +
    ylab("Potential excess antibiotic orders saved (%)") +
    ggtitle(paste0(paste(str_to_title(antibiotic24h), collapse="/"), " -> ", model$antibiotic)) +
    scale_x_reverse() + theme_classic()
    # Hide xlabel text
    
}
```

# Download and prep the data for modeling
## Train: 2009-2021
## Test: 2022, 2023
```{r}
# Download the data from BigQuery
query = "SELECT * FROM `som-nero-phi-jonc101.mvm_abx.encounter_culture_labels_features_inpatient_24h_with_inference`"
q = bq_project_query("som-nero-phi-jonc101", query)
df = bq_table_download(q, page_size = 10000)

# Add Charlson comorbidity index scores
df = df %>% add_charlson_comorbidity_index("csn")

# Clean data frame (remove missing labels, set features types, etc)
df = df %>% 
  filter(!is.na(resistance)) %>% # Remove rows with missing resistance label
  filter(year != "2024") %>% # Remove rows in year 2024
  filter(age >= 18) %>% # Only adult patients
  mutate(across(c("antibiotic", all_of(categorical_features)), as.factor)) %>%
  mutate(across(all_of(numeric_features), as.numeric)) %>%
  unique()

# Feature list
print(c(numeric_features, categorical_features) %>% setdiff(c("year")))
```
# Fit LR model for Cefazolin (best AUC) with imputed data
``` {r}
model_features = c(categorical_features, numeric_features)

---
title: "Inpatient Encounter-Culture Antibiotic Sensitivity Prediction - Stanford only"
author: "Manoj Maddali, MD"
date: "`r Sys.Date()`"
output: html_document
---
  
```{r setup, include=FALSE}
library(tidyverse)
library(bigrquery)
library(gt)
library(caret)
library(xgboost)
library(pROC)
library(MLmetrics)
library(comorbidity)
library(lightgbm)
library(parallel)
library(shapviz)

bq_auth()
```

# Define helper functions
``` {r}
#########################################################################################################################
# Define features for use in model
#########################################################################################################################
categorical_features = c("female", "dialysis", "year", "vasopressor")
numeric_features = c("age", "cci",
                     "wbch", "hgbl", "pltl",
                     "sodiuml", "bicarbl", "bunh", "creath", "glucl",
                     "asth", "alth", "bilih",
                     "albuml",  "lactateh",
                     "hrateh", "rrateh", "temph", "sysbpl", "diasbpl", "weightl",
                     "count_rx_oral", "count_rx_iv", "number_of_admissions",
                     "count_ciprofloxacin_resistance_within_6mo", "count_ciprofloxacin_resistance_before_6mo",
                     "count_oxacillin_resistance_within_6mo", "count_oxacillin_resistance_before_6mo",
                     "count_ampicillin_resistance_within_6mo", "count_ampicillin_resistance_before_6mo",
                     "count_cefazolin_resistance_within_6mo", "count_cefazolin_resistance_before_6mo",
                     "count_ceftriaxone_resistance_within_6mo", "count_ceftriaxone_resistance_before_6mo",
                     "count_cefepime_resistance_within_6mo", "count_cefepime_resistance_before_6mo",
                     "count_pip_tazo_resistance_within_6mo", "count_pip_tazo_resistance_before_6mo",
                     "count_meropenem_resistance_within_6mo", "count_meropenem_resistance_before_6mo",
                     "count_trim_sulfa_resistance_within_6mo", "count_trim_sulfa_resistance_before_6mo",
                     "count_vancomycin_resistance_within_6mo", "count_vancomycin_resistance_before_6mo",
                     "prior_year_antibiogram_ciprofloxacin",
                     "prior_year_antibiogram_oxacillin",
                     "prior_year_antibiogram_ampicillin",
                     "prior_year_antibiogram_cefazolin",
                     "prior_year_antibiogram_ceftriaxone",
                     "prior_year_antibiogram_cefepime",
                     "prior_year_antibiogram_pip_tazo",
                     "prior_year_antibiogram_meropenem",
                     "prior_year_antibiogram_trim_sulfa",
                     "prior_year_antibiogram_vancomycin",
                     "count_prior_pseudomonas_culture",
                     "count_prior_staph_aureus_culture",
                     "count_prior_enterococcus_culture")

#########################################################################################################################
# Calculate and add Charlson comorbidity index scores from parsed ICD9 codes
# Using "comorbidity" package in R
#########################################################################################################################
add_charlson_comorbidity_index = function(df, index) {
  # First parse/explode values of icd9s column for each patient
  icds_parsed = df %>% separate_longer_delim(c(icd9s), delim=",")
  
  # Calculate individual component comorbidity scores
  comorbidity_scores = comorbidity(x=icds_parsed, id=index, code="icd9s", map="charlson_icd9_quan", assign0=TRUE, tidy.codes=TRUE)

  # Score the comorbidity index (unadjusted for age)
  comorbidity_scores$cci = score(comorbidity_scores, assign0 = TRUE)  
  
  # Add the comorbidity index to the original dataframe
  df = df %>% left_join(comorbidity_scores %>% select(all_of(c(index, "cci"))), by=index)
    
  # Adjust the charlson comorbidity index for age
  df = df %>% mutate(cci = cci + ifelse(age >= 80, 4,  
                                        ifelse(age >= 70, 3, 
                                               ifelse(age >= 60, 2, 
                                                      ifelse(age >= 50, 1, 0)))))  
  
  return(df)
}

#########################################################################################################################
# Model training for LightGBM models
#########################################################################################################################
train_test_lgb = function(df, model_antibiotic, features) {
  # Select data for the given antibiotic and features
  df = df %>% filter(antibiotic == model_antibiotic) %>%
    select(csn, resistance, all_of(features)) %>%
    unique()
  
  # Split data into training (pre-2022) and testing (2022 and 2023)
  df_train = df %>% filter(!year %in% c("2022", "2023")) %>% select(-year)
  df_test = df %>% filter(year %in% c("2022", "2023")) %>% select(-year)

  # Set up LightGBM datasets
  df_lgb_train = lgb.Dataset(as.matrix(df_train %>% select(-csn, -resistance)), label = df_train$resistance)
  df_lgb_validation = lgb.Dataset(as.matrix(df_test %>% select(-csn, -resistance)), label = df_test$resistance)

  # Set up hyperparameter grid search
  lgb_grid = expand.grid(
    nrounds = 1000, # early_stopping_rounds will stop training before this if the validation metric doesn't improve
    learning_rate = c(0.01, 0.025, 0.05, 0.1),
    bagging_fraction = c(0.6, 0.8, 0.9, 1),
    bagging_freq = c(5, 10, 20),
    early_stopping_rounds = NA)

  # Set up vector to save each hyperparameter's cross-validated AUC
  cv.aucs = rep(NA, nrow(lgb_grid))

  # Iterate through hyperparameters to find the best cross-validated AUC
  for (i in 1:nrow(lgb_grid)) {
    # Train and test the LightGBM model with current hyperparameters
    model.lgb = lgb.cv(data = df_lgb_train,
                          nrounds = 1000,
                          nfold = 5,
                          params = list(objective = "binary",
                                        metric = "auc",
                                        is_unbalance = TRUE, # Automatically calculate positive weights for unbalanced classes
                                        num_threads = detectCores(), # Use all available cores
                                        learning_rate = lgb_grid[i, "learning_rate"],
                                        max_depth = -1,
                                        bagging_fraction = lgb_grid[i, "bagging_fraction"],
                                        bagging_freq = lgb_grid[i, "bagging_freq"],                                    
                                        early_stopping_rounds = 50,
                                        verbosity = -1))
  
    # Save the AUC
    cv.aucs[i] = model.lgb$best_score
  
    # Save the early_stopping_rounds
    lgb_grid[i, "early_stopping_rounds"] = model.lgb$best_iter
  }

  # Train the final model with the best hyperparameters
  model.lgb = lgb.train(data = df_lgb_train,
                        nrounds = lgb_grid[which.max(cv.aucs), "early_stopping_rounds"],
                        valids = list(validation = df_lgb_validation),
                        params = list(objective = "binary",
                                      metric = "auc",
                                      is_unbalance = TRUE, # Automatically calculate positive weights for unbalanced classes
                                      num_threads = detectCores(), # Use all available cores
                                      learning_rate = lgb_grid[which.max(cv.aucs), "learning_rate"],
                                      max_depth = -1,
                                      bagging_fraction = lgb_grid[which.max(cv.aucs), "bagging_fraction"],
                                      bagging_freq = lgb_grid[which.max(cv.aucs), "bagging_freq"],                                    
                                      verbosity = -1))
  
  # Make predictions
  predictions = predict(model.lgb, as.matrix(df_test %>% select(-csn, -resistance)))
  
  # Calculate ROC
  roc = roc(predictor = predictions, df_test$resistance) 
  
  # Generate confusion matrix
  cm = confusionMatrix(factor(ifelse(predictions >= 0.5, 1, 0)),
                     factor(df_test$resistance), positive = "1", mode = "everything")
  
  # Return the model, AUC, ROC, and confusion matrix
  return(list(antibiotic = str_to_title(model_antibiotic),
              model = model.lgb,
              best_tune = lgb_grid[which.max(cv.aucs),],
              df_train = df_train,
              df_test = df_test,
              predictions = predictions,
              auc = roc$auc,
              roc = roc,
              cm = cm))
}

#########################################################################################################################
# Calculate Shapeley values and plot feature importance
#########################################################################################################################
plot_shap_importance = function(model) {
  shap = shapviz(model$model,
                 X=as.matrix(model$df_train %>% select(-csn, -resistance)),
                 X_pred=as.matrix(model$df_train %>% select(-csn, -resistance)))
  
  shap %>%
    sv_importance(kind="beeswarm") + ggtitle(paste0("SHAP Feature Analysis for ", model$antibiotic, " Model"))
}

#########################################################################################################################
# Plot antibiotic orders saved in next 24 hours
#########################################################################################################################
antibiotics_24h_saved = function(model, df_all_antibiotics_24h, antibiotic24h) {
  # Load test data and predictions from model
  df_test = model$df_test
  df_test$predictions = model$predictions
  
  # Merge with 24h antibiotic order df with df_test
  df_all_antibiotics_24h = df_test %>%
    left_join(df_all_antibiotics_24h %>% select(csn, antibiotic_24h) %>% unique(), by="csn")

  # Filter out the 24h antibiotic of interest
  if (length(antibiotic24h) > 2) {
    stop("Antibiotic24h must only contain max 2 antibiotics")
  }
  
  # Filter out the 24h antibiotic of interest
  if (length(antibiotic24h) == 1) {
    specific_antibiotics_24h = df_all_antibiotics_24h %>%
      filter(antibiotic_24h == str_to_upper(antibiotic24h[1])) %>%
      select(-antibiotic_24h) %>%
      unique()
  } else { # If there are two antibiotics, filter out the common CSNs and take the set of CSNs with both antibiotics
    csn1 = df_all_antibiotics_24h %>% filter(antibiotic_24h == str_to_upper(antibiotic24h[1])) %>% pull(csn)
    csn2 = df_all_antibiotics_24h %>% filter(antibiotic_24h == str_to_upper(antibiotic24h[2])) %>% pull(csn)    
    csns = intersect(csn1, csn2)
    
    specific_antibiotics_24h = df_all_antibiotics_24h %>%
      filter(csn %in% csns) %>%
      select(-antibiotic_24h) %>%
      unique()
  }
  
  # Create a sequence of thresholds to test NPV
  seq = seq(0.00, 1, 0.01)
  excess = data.frame(matrix(ncol = 6, nrow = length(seq)))
  colnames(excess) = c("threshold", "npv", "excess", "relative_excess", "true_excess", "perc_saved")
  excess$threshold = seq
  
  # True excess is the number of orders that were not resistant on culture (gold standard)
  excess$true_excess = nrow(specific_antibiotics_24h %>% filter(resistance == 0))

  # Iterate through thresholds and calculate NPV and excess antibiotic prescriptions
  for (i in seq) {
    # Predict resistance based on threshold
    specific_antibiotics_24h = specific_antibiotics_24h %>% mutate(predicted_resistance = ifelse(predictions >= i, 1, 0))
    # Calculate confusion matrix for NPV
    cm = confusionMatrix(factor(specific_antibiotics_24h$predicted_resistance),
                         factor(specific_antibiotics_24h$resistance),
                         positive = "1", mode = "everything")
    excess[excess$threshold == i, "npv"] = cm$byClass['Neg Pred Value']
    # Calculate number of excess antibiotic orders (predicted resistance = 0, true resistance = 0)
    excess[excess$threshold == i, "excess"] = nrow(specific_antibiotics_24h %>% filter(predicted_resistance == 0, resistance == 0))
    # Calculate relative excess (true excess - excess)
    excess[excess$threshold == i, "relative_excess"] = excess[excess$threshold == i, "true_excess"] - excess[excess$threshold == i, "excess"]
    # Calculate percentage of excess antibiotic orders saved
    excess[excess$threshold == i, "perc_saved"] = round(100*excess[excess$threshold == i, "excess"] / excess[excess$threshold == i, "true_excess"], 2)
  }
  
  # Plot excess antibiotic orders saved vs NPV
  excess %>%
    group_by(npv_group = cut(npv, breaks = seq(0, 1, 0.025))) %>%
    summarize(mean_npv = mean(npv), mean_perc_saved = mean(perc_saved)) %>%
    filter(!is.na(mean_npv)) %>%
    filter(mean_npv > 0.7) %>% 
    ggplot() +
    geom_line(aes(x=mean_npv, y=mean_perc_saved)) +
    geom_vline(xintercept=0.8, linetype="dotted") + 
    geom_ribbon(aes(x=mean_npv, ymin=0, ymax=mean_perc_saved), fill="red", alpha=0.2) +
    xlab("Negative predictive value") +
    ylab("Potential excess antibiotic orders saved (%)") +
    ggtitle(paste0(paste(str_to_title(antibiotic24h), collapse="/"), " -> ", model$antibiotic)) +
    scale_x_reverse() + theme_classic()
    # Hide xlabel text
    
}
```

# Download and prep the data for modeling
## Train: 2009-2021
## Test: 2022, 2023
```{r}
# Download the data from BigQuery
query = "SELECT * FROM `som-nero-phi-jonc101.mvm_abx.encounter_culture_labels_features_inpatient_24h_with_inference`"
q = bq_project_query("som-nero-phi-jonc101", query)
df = bq_table_download(q, page_size = 10000)

# Add Charlson comorbidity index scores
df = df %>% add_charlson_comorbidity_index("csn")

# Clean data frame (remove missing labels, set features types, etc)
df = df %>% 
  filter(!is.na(resistance)) %>% # Remove rows with missing resistance label
  filter(year != "2024") %>% # Remove rows in year 2024
  filter(age >= 18) %>% # Only adult patients
  mutate(across(c("antibiotic", all_of(categorical_features)), as.factor)) %>%
  mutate(across(all_of(numeric_features), as.numeric)) %>%
  unique()

# Feature list
print(c(numeric_features, categorical_features) %>% setdiff(c("year")))
```
# Cefazolin logistic regression model with imputed data
``` {r}
model_features = c(categorical_features, numeric_features)
  
# Select data for the given antibiotic and features
df = df %>% filter(antibiotic == "CEFAZOLIN") %>%
  select(csn, resistance, all_of(model_features)) %>%
  unique()

df_imputed = mice(df, method = "rf", seed = 123) %>% complete()

# Split the data into train and test sets
set.seed(123)
df_train_imputed = df_imputed %>% filter(!year %in% c("2022", "2023")) %>% select(-year, -csn)
df_test_imputed = df_imputed %>% filter(year %in% c("2022", "2023")) %>% select(-year, -csn)

# Logistic regression model
model = glm(resistance ~ ., data = df_train_imputed, family = "binomial")

# Predict on the test set
predictions = predict(model, newdata = df_test_imputed, type = "response")

# Calculate AUC
roc = roc(df_test_imputed$resistance, predictions)
roc$auc
```
