Matching: to balance two cohorts of patients based a set of given covariates, minimizing potentially confounding biases

Parameters:
1) cohortA_file: absolute or relative path to csv file containing cohort A dataset
2) cohortB_file: absolute or relative path to csv file containing cohort B dataset
3) labelA: a label/string to describe the patients in cohort A
4) labelB: a label/string to describe the patients in cohort B
5) id_column: integer corresponding to the ID colummn of the patient
6) binary_columns: c-vector (e.g. c(1,2,3)) corresponding to column indices for binary columns
7) covariate_columns: c-vector corresponding to covariates to conduct matching on
8) custom_formatting: boolean that enables the user to add custom formatting code in a specially allocated slot in the code below; 1 = custom formatting is turned on, 0 = custom formatting is turned off
9) overlap_plot: boolean that enables the user to plot an overlap of propensity scores between the two patient cohorts; 1 = plot, 0 = don't plot
10) matched_output_file: absolute or relative path to csv file to which the matched dataset will be written to

```{r}
match_cohorts <- function(cohortA_file, cohortB_file, labelA, labelB, id_column, binary_columns, covariate_columns, custom_formatting,
                          overlap_plot, matched_output_file) {
  
  # Load in required libraries
  library(caret)
  library(ggplot2)
  library(dplyr)
  library(mice)
  library(Matching)
  library(tableone)
  require(reshape2)
  Sys.setenv(TZ='UTC')  
    
  ###################################
  ### Read and format data frames ###
  ###################################
  print("Reading and formatting data frames...")
    
  # Read tab-delimited input using header row labels and interpret Python None strings as R NA values
  cohortA_df = tbl_df(read.csv(cohortA_file,header=TRUE,sep=",",na.strings="None", stringsAsFactors=FALSE));
  cohortB_df = tbl_df(read.csv(cohortB_file,header=TRUE,sep=",",na.strings="None",stringsAsFactors=FALSE));
  
  # Add labels: 1 = labelA, 0 = labelB
  cohortA_df = data.frame(cohortA_df, label=labelA)
  print("Dimensions of cohort A:")
  dim(cohortA_df)
  cohortB_df = data.frame(cohortB_df, label=labelB)
  print("Dimensions of cohort B:")
  dim(cohortB_df)
  label_column <- dim(cohortA_df)[2]
  
  # Combine cohorts into one dataframe
  df_combined <- rbind(cohortA_df, cohortB_df)
  
  # Convert T/F = 1/0 (Columns labeled under binary_columns)
  for (col_id in binary_columns) {
    df_combined[,col_id] <- factor(ifelse(df_combined[,col_id]=="True", 1, 0))
  }
  # Condense dataframe
  columns_of_interest = c(id_column, covariate_columns, label_column)
  df_combined <- df_combined[,columns_of_interest]
  
  # Ensure legitimate variable names
  colnames(df_combined) <- make.names(colnames(df_combined))
  
  #########################
  ### Custom Formatting ###
  #########################
  
  # Editable region for you to add additional formatting commands
  if (custom_formatting == "1") { 
    # Remove features with too much NA/missing data (by manual inspection)
    df_combined = df_combined[,c(1:3,6:48)]
  
    # Remove lab tests to reduce bias: columns 7-15 in df
    df_combined = df_combined[,c(1:6, 16:46)]
  }
  
  # Impute the remaining NA values: predictive mean imputation
  md.pattern(df_combined);
  md_data <- mice(df_combined,m=1,maxit=5,meth='pmm',seed=500);
  df <- complete(md_data,1);
  
  ##########################
  ### Propensity Scoring ###
  ##########################
  
  print("Building propensity score model...")
  # Logistic Regression
  label_column <- dim(df)[2]
  label <- factor(ifelse(df[,label_column]==labelA, 1, 0)) 
  glm.fit <- glm(label ~.,family=binomial(link='logit'),data=df[,c(-id_column, -label_column)])
  
  # Propensity Estimation
  cohortA_features <- df %>%
    filter(label == labelA)
  cohortA_features <- cohortA_features[,c(-id_column, -label_column)] # remove id, label
  
  cohortB_features <- df %>%
    filter(label == labelB)
  cohortB_features <- cohortB_features[,c(-id_column, -label_column)] # remove id, label
  
  pred_A <- predict(glm.fit, newdata=cohortA_features, type="response")
  pred_B <- predict(glm.fit, newdata=cohortB_features, type="response")
  
  ####################
  ### Overlap Plot ###
  ####################
  if (overlap_plot == 1) {
    print("Plotting overlap...")
    # note that logit = log(p/(1-p))
    require(car)
    logit_A <- as.numeric(logit(pred_A))
    logit_B <- as.numeric(logit(pred_B))
    histdata <- data.frame(rbind.data.frame(cbind.data.frame(logit=logit_A,label=labelA), cbind.data.frame(logit=logit_B,label=labelB)))
    plot <- ggplot(histdata, aes(x=logit, fill=label)) + geom_histogram(alpha=0.2, position="identity")
    print(plot)
  }
  
  ########################
  ### Caliper Matching ###
  ########################
  print("Conducting caliper matching...")
  # caliper = the maximum tolerated difference between matched subjects in a "non-perfect" matching 
  # conduct matching w/o replacement on logit of propensity
  temp <- data.frame(rbind.data.frame(cbind.data.frame(logit=logit_A,label=1), cbind.data.frame(logit=logit_B,label=0)))
  
  # compute caliper: 0.2*sd(logit(propensity_scores))
  propensity_scores <- temp$logit
  opt_caliper <- 0.2*sd(propensity_scores)
  print(opt_caliper)
  
  #matches <- Match(X=temp$logit, Tr=temp$label, caliper=0.25, replace=FALSE, M=2)
  matches <- Match(X=temp$logit, Tr=temp$label, caliper=opt_caliper, replace=F, M=1)
  
  # number of patients in matched set
  treated <- unique(matches$index.treated)
  control <- unique(matches$index.control)
  print("Post-Matching Cohort Sizes:")
  length(treated)
  length(control)
  
  ##############################
  ### Post-Matching Analysis ###
  ##############################
  
  print("Preparing post-matching results...")

  # Examine Matching Results
  cohortA_full <- df %>% 
    filter(label == labelA)
  
  cohortB_full <- df %>% 
    filter(label == labelB)
  
  df_full <- rbind.data.frame(cohortA_full, cohortB_full)
  # treated = labelA (1), control = labelB (0)
  df_matched <- rbind.data.frame(df_full[treated,], df_full[control,])
  
  # Output to CSV
  write.csv(df_matched, file = matched_output_file)
  
  # Before and After Comparisons
  cohortA_matched <- df_matched %>%
    filter(label == labelA)
  cohortA_matched <- cohortA_matched[,c(-1,-37)]
  cohortB_matched <- df_matched %>%
    filter(label == labelB)
  cohortB_matched <- cohortB_matched[,c(-1,-37)]
  
  cohortA_unmatched <-cohortA_features
  cohortB_unmatched <- cohortB_features
  
  # Pre and Post Matching Standardized Difference in Means
  print("Plotting pre- vs. post-matching SMD values...")
  
  t1 <- cbind(cohortA_unmatched, label=labelA)
  t2 <- cbind(cohortB_unmatched, label=labelB)
  unmatched <- rbind(t1, t2)
  unmatched_tab <- CreateTableOne(vars = colnames(unmatched), strata = "label", data = unmatched, test = FALSE)
  print(unmatched_tab, smd = TRUE)
  
  t1 <- cbind(cohortA_matched, label=labelA)
  t2 <- cbind(cohortB_matched, label=labelB)
  matched <- rbind(t1, t2)
  matched_tab <- CreateTableOne(vars = colnames(matched), strata = "label", data = matched, test = FALSE)
  print(matched_tab, smd = TRUE)
  
  ## Construct a data frame containing variable name and SMD from all methods
  data_plot <- data.frame(variable  = names(ExtractSmd(unmatched_tab)),
                         Unmatched = ExtractSmd(unmatched_tab),
                         Matched   = ExtractSmd(matched_tab))
  
  ## Create long-format data for ggplot2
  data_plot_melt <- melt(data = data_plot,
                       id.vars = c("variable"),
                       variable.name = "Method",
                       value.name = "SMD")
  
  ## Order variable names by magnitude of SMD
  varNames <- as.character(data_plot$variable)[order(data_plot$Unmatched)]
  
  ## Order factor levels in the same order
  data_plot_melt$variable <- factor(data_plot_melt$variable,
                                  levels = varNames)
  
  ## Plot using ggplot2
  plot <- ggplot(data = data_plot_melt[,c(-72)], mapping = aes(x = variable, y = SMD,
                                            group = Method, color = Method)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 0.1, color = "black", size = 0.1) +
  coord_flip() +
  theme_bw() + theme(legend.key = element_blank())
  print(plot)
  
  # Pre and Post-Matching P-Values
  print("Printing pre- vs. post-matching p-values...")
  
  for (i in 1:length(cohortA_features)) {
    print(colnames(cohortA_matched)[i])
    
    if (length(unique(cohortA_matched[,i])) > 2) { #mean
      print("Unmatched:")
      print(labelA)
      print(mean(cohortA_unmatched[,i]))
      #print(sd(cohortA_unmatched[,i]))
      
      print(labelB)
      print(mean(cohortB_unmatched[,i]))
      #print(sd(cohortB_unmatched[,i]))
      #print(wilcox.test(expert_unmatched[,i], everyone_unmatched[,i], alternative = "two.sided")$p.value)
      print("P-value:")
      print(t.test(cohortA_unmatched[,i], cohortB_unmatched[,i])$p.value)
      
      print("Matched:")
      print(labelA)
      print(mean(cohortA_matched[,i]))
      #print(sd(cohortA_matched[,i]))
      
      print(labelB)
      print(mean(cohortB_matched[,i]))
      #print(sd(cohortB_matched[,i]))
      #print(wilcox.test(expert_matched[,i], everyone_matched[,i], alternative = "two.sided")$p.value)
      
      print("P-value:")
      print(t.test(cohortA_matched[,i], cohortB_matched[,i])$p.value)
    }
    else { # fraction    
      print("Unmatched:")
      print(labelA)
      print(length(cohortA_unmatched[,i][cohortA_unmatched[,i] == 1])/dim(cohortA_unmatched)[1])
      
      print(labelB)
      print(length(cohortB_unmatched[,i][cohortB_unmatched[,i] == 1])/dim(cohortB_unmatched)[1])
      #print(chisq.test(table(expert_unmatched[,i], everyone_unmatched[,i]))$p.value)
      
      print("P-value:")
      print(t.test(as.numeric(cohortA_unmatched[,i]), as.numeric(cohortB_unmatched[,i]))$p.value)
  
      print("Matched:")
      print(labelA)
      print(length(cohortA_matched[,i][cohortA_matched[,i] == 1])/dim(cohortA_matched)[1])
      
      print(labelB)
      print(length(cohortB_matched[,i][cohortB_matched[,i] == 1])/dim(cohortB_matched)[1])
      #print(chisq.test(table(expert_matched[,i], everyone_matched[,i]))$p.value)
      
      print("P-value:")
      print(t.test(as.numeric(cohortA_matched[,i]), as.numeric(cohortB_matched[,i]))$p.value)
    }
  }
}
```

```{r}
#############################
### Example Function Call ###
#############################

# Set working directory if necessary
setwd("/Users/jwang/Desktop/Patient Matching/")

# Set parameters
cohortA_file <- "expert_2013only.csv"
cohortB_file <- "everyone_2013only.csv"
labelA <- "expert"
labelB <- "everyone"
id_column <- 1
binary_columns <- c(22:49)
covariate_columns <- c(6:51)
custom_formatting <- 1
overlap_plot <- 1
matched_output_file <- "matched.csv"

match_cohorts(cohortA_file, cohortB_file, labelA, labelB, id_column, binary_columns, covariate_columns, custom_formatting, overlap_plot, matched_output_file)
```