---
title: 'TidyVerse: Python and R'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
mtcars2 <- mtcars 
colnames(mtcars2)[1] <- "car"
write.csv(mtcars2, "mtcars2.csv")
```


## Reading CSV in R versus Python 

#### Chunks are useful for interactively writing code and publishing documents to share 
#### 

```{python}
import pandas as pd
df = pd.read_csv('mtcars2.csv')
print df
```


```{r, echo=FALSE}
library(tidyverse)
df <- read.csv('mtcars2.csv')
df
```

# Filtering Data in R


```{r}
#filter(df, mpg > 15 & mpg < 30) 
#filter(df, car == 'Toyota Corolla') 
#df %>% filter(car != 'Toyota Corolla')
```

# EXERCISE 

1) Filter for column wt greater than 2 

```{r}
# INSERT CODE HERE 
```


# Filtering Data in Python 

```{python}
import pandas as pd
df = pd.read_csv('mtcars2.csv')
#print df
#print df[(df['mpg'] > 15) & (df['mpg'] < 30)]
#print df[df['car'] == 'Toyota Corolla']
#print df[df['car'] != 'Toyota Corolla']

```

# Piping Data in R 

```{r}
iris %>%
  subset(Sepal.Length > 5) %>%
  aggregate(. ~ Species, ., mean)
```

# BEFORE PIPES 

```{r}
a <- filter(mtcars, carb > 1)

b <- group_by(a, cyl)

c <- summarise(b, Avg_mpg = mean(mpg))

d <- arrange(c, Avg_mpg)

e <- arrange(c, desc(Avg_mpg))

a 
b
c
d
e
```

```{r}
mtcars %>%
        filter(carb > 1) %>%
        group_by(cyl) %>%
        summarise(Avg_mpg = mean(mpg)) %>%
        arrange(desc(Avg_mpg))
```

## EXERCISE 2 

2) Write your own piped where you filter hp greater than 100, Group by carb, summarize the average mpg in 
  ascending order 
  
```{r}
# INSERT CODE HERE 
```
  

## Bonus Interactive Visualization in 1 line! 

```{r}
library(plotly)
p <- ggplot(mtcars, aes(wt, mpg))
p1 <- p + geom_point(size=4)
p2 <- p1 + geom_point(aes(colour=factor(cyl)))
p2
```

```{r}
ggplotly(p2)
```

