---
title: "Classifier Performance"
author: "Rachael Caelie Aikens"
date: "May 1, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = 'data')

library(ggplot2)
library(gridExtra)
library(data.table)
library(knitr)
library(tidyr)
library(dplyr)
library(grid)
library(ggpubr)
library(RColorBrewer)
library(scales)
```

# Functions for loading and preprocessing data

```{r basic parameters of analysis}
labs_done_percent <- c("TNI", "CK", "PTT", "PHOS", "MGN", "LACWB", "HEPAR", "CRP", "K", "FER", "PLTS", "HCTX", "LIPS", "NTBNP", "TSH")
labs_done_sd <- c("A1C", "CK", "TNI", "PTT", "PHOS", "MGN", "HEPAR", "CRP",  "FER", "PLTS", "LIPS", "NTBNP", "TSH", "USPG", "LACWB")

labs_done_percent <- intersect(labs_done_percent, labs_done_sd)
labs_done_sd <- intersect(labs_done_percent, labs_done_sd)

full_lab_names <- paste("LAB", union(labs_done_percent, labs_done_sd), sep = "")

# upload Component map
component_map <- read.csv("../LabResultMap.tab", sep = "\t") %>%
  filter(proc_code %in% full_lab_names) %>%
  group_by(proc_code) %>%
  summarize(description = first(description))

# rename to lower case
component_map$description <- c("Creatinine kinase", "C-reactive protein", "Ferritin", "Heparin Activity",
                               "Lactate, whole blood", "Lipase", "Magnesium, serum/plasma", "NT-PROBNP",
                               "Phospherous serum/plasma", "Platelet count", "Partial Throboplastin Time",
                               "Troponin I", "Thyroid Stimulating Hormone")

params <- 1:5*10
n_models <- 6
```

```{r functions to load data}
# construct from filepath and load
load_data <- function(lab, params = 1:5*10, n_models = 6, change_def = "percent"){
  paths <- paste("LAB", lab, "/change_", change_def,"_0", params/10, "/LAB", lab, "-change-prediction-report.tab",
               sep = "")
  report <- rbindlist(lapply(paths, read.csv,
                             sep = "\t",
                             comment.char = "#",
                             stringsAsFactors = FALSE))
  return(report)
}
```

```{r preprocessing functions}
# parse dictionary string and return number of cases
get_n_cases <- function(y_dict){
  str_list <- gsub('.{1}$', '', strsplit(y_dict, " ")[[1]])
  case <- as.numeric(str_list[4])
  return(case)
}

# parse dictionary string and return number of control
get_n_controls <- function(y_dict){
  str_list <- gsub('.{1}$', '', strsplit(y_dict, " ")[[1]])
  control <- as.numeric(str_list[2])
  return(control)
}

# preprocess raw loaded data
preprocess_data <- function(report, lab, params = 1:5*10, n_models = 6, comp_map = component_map){
  report$change_param <-  sort(rep(params, each = n_models))
  report$LAB <- lab
  report$test_cases <- sapply(report$y_test.value_counts..,
                                        function(x) get_n_cases(x))
  report$test_controls <- sapply(report$y_test.value_counts..,
                                        function(x) get_n_controls(x))
  report$proportion_unchanged <- report$test_cases/(report$test_cases + report$test_controls)
  report$proc_code <- paste("LAB", report$LAB, sep = "")
  result <- left_join(report, comp_map, by = "proc_code") %>% select(-proc_code)
  return(result)
}

# run load and preprocessing and return processed result data
load_and_process <- function(lab, params = 1:5*10, n_models = 6, change_def = "percent"){
  raw <- load_data(lab, params, n_models, change_def)
  return(preprocess_data(raw, lab, params, n_models))
}
```

# Analysis

## Many Labs

```{r}
full_data_percent <- rbindlist(lapply(labs_done_percent, load_and_process, params = params, change_def = "percent"))
full_data_sd <- rbindlist(lapply(labs_done_sd, load_and_process, params = params, change_def = "sd")) %>%
  mutate(change_param = change_param/100)
```

```{r}
summarize_group_report <- function(full_data){
  result <- full_data %>%
  group_by(change_param, LAB) %>%
  summarize(max_roc = max(roc_auc),
            max_roc_lowCI = nth(roc_auc_0.95_lower_ci, which.max(roc_auc)),
            max_roc_highCI = nth(roc_auc_0.95_upper_ci, which.max(roc_auc)),
            max_percent_predictability = max(percent_predictably_positive),
            max_pp_lowCI = nth(percent_predictably_positive_0.95_lower_ci,
                               which.max(percent_predictably_positive)),
            max_pp_highCI = nth(percent_predictably_positive_0.95_upper_ci,
                                which.max(percent_predictably_positive)),
            test_cases = first(test_cases),
            test_controls = first(test_controls),
            proportion_unchanged = first(proportion_unchanged),
            description = first(description))

  return(result)
}

report_summary_percent <- summarize_group_report(full_data_percent)
report_summary_sd <- summarize_group_report(full_data_sd)
```

First, we can check how many lab tests are unchanged from the previous measurement, for any given percent threshold.

```{r}
F1A <- ggplot(report_summary_percent %>% filter(change_param == 10) ,
       aes(description, proportion_unchanged)) +
  geom_col(fill = "steelblue3") +
  labs(y = "Proportion of repeat tests \nwithin 10% of previous measurement")+
  coord_flip() +
  ylim(c(0,1))

F1B <- ggplot(report_summary_sd %>% filter(change_param == 0.1) ,
       aes(reorder(description, proportion_unchanged), proportion_unchanged)) +
  geom_col(fill = "steelblue3") +
  labs(y = "Proportion of repeat tests \nwithin 0.1 standard deviations of previous measurement")+
  coord_flip() +
  ylim(c(0,1))

grid.arrange(F1A, F1B, ncol =1)
```

We can combine those:

```{r, fig.height = 3.3in, fig.width = 6.1in}
method_col <- c(rep("± 0.1 SD", dim(report_summary_sd)[1]),
                rep("± 10%", dim(report_summary_percent)[1]))

F1_plot_data <- rbind(report_summary_sd, report_summary_percent) %>%
  ungroup() %>%
  mutate(metric = method_col) %>%
  filter(change_param %in% c(0.1, 10))

F1 <- ggplot(F1_plot_data,
             aes(description, proportion_unchanged)) +
  geom_col(aes(fill = metric), position = "dodge") +
  scale_fill_brewer(palette = "Set1") +
  labs(y = "Proportion of repeat tests \nwithin margin of previous measurement", x = NULL, fill = "Margin\nof Change")+
  coord_flip() +
  ylim(c(0,1)) + theme_bw()

F1
```

Plotting the AU-ROC for the prediction tasks for all labs is too busy. To start, we make a table of our performance just at the 0.1 SD or 10% change level.

```{r}
kable(F1_plot_data %>% select(LAB, metric, max_roc, max_percent_predictability), digits = 2)
```



Next, we filter our results to include only the lab tests for which the percent predictably positive we achieved a lower CI greater than 10 percent for at least one definition of "change."

```{r filter results}
Labs_to_keep_percent <- report_summary_percent %>%
  group_by(LAB) %>%
  summarise(pp_CI_lower = max(max_pp_lowCI)) %>%
  filter(pp_CI_lower >= 0.1)

Labs_to_keep_percent

# proportion of the labs which make this threshold
dim(Labs_to_keep_percent)[1]/length(labs_done_percent)

report_summary_percent_best <- filter(report_summary_percent,
                              LAB %in% Labs_to_keep_percent$LAB)
```


```{r}
auroc_trend_plot_bygroup <- function(report_summary, my_xlab = "Change Threshold"){
  myplot <- ggplot(report_summary,
         aes(x = change_param, y = max_roc, group = LAB, color = LAB)) +
    geom_line()+
    labs(x = my_xlab, y = "AU-ROC")+
    ylim(c(0.5,1))
  return(myplot)
}

F3A <- auroc_trend_plot_bygroup(report_summary_percent_best, "Percent change threshold")
```

```{r, figure.width = 3}
ppp_trend_plot_bygroup <- function(report_summary, my_xlab = "Change Threshold"){
  myplot <- ggplot(report_summary,
         aes(x = change_param, y = max_percent_predictability,
             group = LAB, color = LAB)) +
    geom_line() +
    labs(x = my_xlab, y = "Percent predictibly positive")+
    ylim(c(0,1))
  return(myplot)
}

F3B <- ppp_trend_plot_bygroup(report_summary_percent_best, "Percent change threshold")
```

```{r, fig.height = 4.5in, fig.width = 7.25}
ppp_bar_plot <- function(report_summary, fill_label = "Change Threshold", mypalette = "Reds"){
  myplot <- ggplot(report_summary, aes(description, max_percent_predictability))+
    geom_col(aes(fill = as.factor(change_param)), position = "dodge") +
    labs(fill = fill_label, y = "Percent Predictability@99%", x = NULL) +
    coord_flip() +
    scale_fill_brewer(palette = mypalette, direction= -1) +
    ylim(c(0,1)) + theme_bw() #+
#    theme(legend.key.size = unit(0.25, "cm"),
#          legend.title=element_text(size=8),
#          legend.text=element_text(size=8))
  return(myplot)
}

roc_bar_plot <- function(report_summary, fill_label = "Change Threshold", mypalette = "Reds"){
  myplot <- ggplot(report_summary, aes(description, max_roc))+
    geom_col(aes(fill = as.factor(change_param)), position = "dodge") +
    labs(fill = fill_label, y = "AU-ROC", x = NULL) +
    coord_flip() +
    scale_y_continuous(limits = c(0.5, 1), oob = rescale_none) +
    scale_fill_brewer(palette = mypalette, direction= -1) + theme_bw() #+
#   theme(legend.key.size = unit(0.25, "cm"),
#          legend.title=element_text(size=8),
#          legend.text=element_text(size=8))
  return(myplot)
}

F2B <- ppp_bar_plot(report_summary_percent, "Percent change threshold", "Blues")
F2A <- roc_bar_plot(report_summary_percent, "Percent change threshold", "Blues")
ggarrange(F2A, F2B, common.legend = TRUE, legend = "bottom", labels = "AUTO")
F3B <- ppp_bar_plot(report_summary_sd, "SD change threshold", "Reds")
F3A <- roc_bar_plot(report_summary_sd, "SD change threshold", "Reds")
ggarrange(F3A, F3B, common.legend = TRUE, legend = "bottom", labels = "AUTO")
```

```{r}
Labs_to_keep_sd <- report_summary_sd %>%
  group_by(LAB) %>%
  summarise(pp_CI_lower = max(max_pp_lowCI)) %>%
  filter(pp_CI_lower >= 0.1)

Labs_to_keep_sd

# proportion of the labs which make this threshold
dim(Labs_to_keep_sd)[1]/length(labs_done_sd)

report_summary_sd_best <- filter(report_summary_sd,
                              LAB %in% Labs_to_keep_sd$LAB)
```

```{r}
F3A <- auroc_trend_plot_bygroup(report_summary_sd_best, "Standard deviation change threshold")
F3B <- ppp_trend_plot_bygroup(report_summary_sd_best, "Standard deviation change threshold")

ggarrange(F3A, F3B, ncol = 2, common.legend = TRUE, legend = "bottom")
```


```{r, eval = FALSE, include = FALSE, echo = FALSE}
# this plot isn't terribly informative
ggplot(report_summary_best, aes(x = percent_change, y = test_cases/(test_cases+test_controls), group = LAB, color = LAB)) +
  geom_line() +
  #geom_line(aes(x = percent_change, y = test_cases/(test_cases+test_controls)), linetype = "dashed") +
  #geom_line(aes(x = percent_change, y = max_percent_predictability*test_cases/(test_cases+test_controls)), linetype = "dashed")+
  labs(x = "Percent change threshold", y = "Percent with label 1")+
  ylim(c(0,1))
```

## Save Results to a File

```{r}
# all raw data
write.table(full_data_percent, file = "full_data_percent.tsv", sep = "\t", quote = FALSE, row.names = FALSE)
write.table(full_data_sd, file = "full_data_sd.tsv", sep = "\t", quote = FALSE, row.names = FALSE)

# processed data
write.table(report_summary_percent, file = "processed_data_percent.tsv", sep = "\t", quote = FALSE, row.names = FALSE)
write.table(report_summary_sd, file = "processed_data_sd.tsv", sep = "\t", quote = FALSE, row.names = FALSE)
```

# Volume Analaysis

```{r upload volume info}
volume_charge_map <- read.csv("../repeated_labs_charges_volumes.csv") %>%
  filter(name %in% full_lab_names,
         window == 7)

report_summary_percent$name <- paste("LAB", report_summary_percent$LAB, sep = "")
report_summary_sd$name <- paste("LAB", report_summary_sd$LAB, sep = "")
vol_charge_table_percent <- left_join(report_summary_percent, volume_charge_map, by = "name") %>% select(-name)
vol_charge_table_sd <- left_join(report_summary_sd, volume_charge_map, by = "name") %>% select(-name)
```

```{r, fig.height = 4.5in, fig.width = 7.25}
predictable_vol_bar_plot <- function(report_summary, fill_label = "Change Threshold", mypalette = "Reds"){
  myplot <- ggplot(report_summary, aes(description, max_percent_predictability*num_repeated, fill = as.factor(change_param))) +
    geom_bar(position = "dodge", stat = "identity") +
    labs(fill = fill_label, y = "Predictable volume (log scale)", x = NULL) +
    geom_text(aes(label=comma(round(signif(max_percent_predictability*num_repeated,2)))),
              hjust = -0.05, position=position_dodge(width=0.9), size = 3) +
    coord_flip() +
    scale_fill_brewer(palette = mypalette, direction= -1) +
    theme_bw() +
    scale_y_log10(labels = comma, limits = c(1, 200000)) +
    theme(legend.key.size = unit(0.3, "cm"),
          legend.title=element_text(size=9),
          legend.text=element_text(size=9),
          axis.title.x = element_text(size = 10),
          axis.text.x = element_text(angle = 45, hjust = 1))
  return(myplot)
}

F4A <- predictable_vol_bar_plot(vol_charge_table_percent %>% filter(change_param <= 20),
                         fill_label = "Percent change\nthreshold", mypalette = "Blues")
F4B <- predictable_vol_bar_plot(vol_charge_table_sd %>% filter(change_param <= 0.2),
                         fill_label = "SD change\nthreshold", mypalette = "Reds")

ggarrange(F4A, F4B, ncol = 2, common.legend = FALSE, legend = "top", labels = "AUTO", hjust = -1)
```

```{r}
options(scipen = 1000000)
predictable_charge_bar_plot <- function(report_summary, fill_label = "Change Threshold", mypalette = "Reds"){
  myplot <- ggplot(report_summary, aes(description, max_percent_predictability*num_repeated*median_price, fill = as.factor(change_param))) +
    geom_col(position = "dodge") +
    labs(fill = fill_label, y = "Predictable charge (log scale)", x = NULL) +
    coord_flip() +
    scale_fill_brewer(palette = mypalette, direction= -1) +
    geom_text(aes(label=paste("$", comma(round(signif(max_percent_predictability*num_repeated*median_price,2))), sep = "")),
              hjust = -0.05, position=position_dodge(width=0.9), size = 2.75) +
    theme_bw() +

    scale_y_log10(labels = scales::dollar, limits = c(1, 1700000000)) +
    theme(legend.key.size = unit(0.3, "cm"),
          legend.title=element_text(size=9),
          legend.text=element_text(size=9),
          axis.title.x = element_text(size = 10),
          axis.text.x = element_text(angle = 45, hjust = 1))
  return(myplot)
}

F5A <- predictable_charge_bar_plot(vol_charge_table_percent %>% filter(change_param <= 20),
                         fill_label = "Percent change\nthreshold", mypalette = "GnBu")
F5B <- predictable_charge_bar_plot(vol_charge_table_sd %>% filter(change_param <= 0.2),
                         fill_label = "SD change\nthreshold", mypalette = "YlGn")

ggarrange(F5A, F5B, ncol = 2, common.legend = FALSE, legend = "top", labels = "AUTO", hjust = -1)

ggarrange(F4A, F4B, F5A, F5B, common.legend = F, legend = "top", labels = "AUTO", hjust = -1, heights = c(1,1.05))
```
