Matching: given patient demographic data, initial vital signs, past medical history, admission diagnosis, treatment team, and other relevant covariates collected between encountering the physician for two cohorts, conduct propensity score matching, yielding two balanced cohorts

PART 1: LOAD IN TOP AND BOTTOM PATIENT COHORTS
```{r}
setwd("/Users/jwang/Desktop/Results/scripts")

# Load in required libraries
library(caret)
library(ggplot2)
library(dplyr)
library(mice)
library(Matching)
library(tableone)
require(reshape2)
require(ISLR)
Sys.setenv(TZ='UTC')  

#######################
### Load Top Cohort ###
#######################
print("Reading and formatting data frames...")
  
# Read tab-delimited input using header row labels and interpret Python None strings as R NA values
df_raw_top = tbl_df(read.csv("/Users/jwang/Desktop/Results/2010-2013_unmatched_patient_feature_matrix_top_performing.csv",header=TRUE,sep=",",na.strings="None", stringsAsFactors=FALSE));

# Convert T/F = 1/0 (Columns labeled under binary_columns)
for (col_id in c(22:61)) {
  df_raw_top[,col_id] <- factor(ifelse(df_raw_top[,col_id]=="True", 1, 0))
}
# Condense dataframe
columns_of_interest = c(1, 6:296) # id, relevant covariates, label columns
df_raw_top <- df_raw_top[,columns_of_interest]

# Ensure legitimate variable names
colnames(df_raw_top) <- make.names(colnames(df_raw_top))

###################
### Missingness ###
###################

# Remove features with too much NA/missing data (by manual inspection)
df_raw_top = df_raw_top[,c(1:3,6:292)]

# Assess lab tests for percentage na: columns 7-15 in df
for (i in 7:15) {
  ratio_na = sum(is.na(df_raw_top[,i]))/dim(df_raw_top)[1]
  print(colnames(df_raw_top)[i])
  print(ratio_na)
}

# Remove urine, column 7 (missingness > 20%)
df_raw_top = df_raw_top[,c(1:6, 8:290)]

# Remove TT and admission diagnosis columns that are all zero
df_raw_top <- data.frame(df_raw_top)
zero_cols = c()
for (i in c(43:54, 57:284)) { # colindex 55 = income, colindex 56 = age
  # replace NA with 0
  df_raw_top[,i][is.na(df_raw_top[,i])] <- 0
  if ((length(df_raw_top[,i][df_raw_top[,i] == 1])) == 0) {
    print(i)
    print(colnames(df_raw_top)[i])
    zero_cols <- c(zero_cols,i)
  }
}
df_raw_top <- df_raw_top[,-zero_cols]

# Convert remaining admission diagnosis columns to factor
for (i in c(56:185)) {
  df_raw_top[,i] = as.factor(df_raw_top[,i])
}

# Impute the remaining NA values: predictive mean imputation
md.pattern(df_raw_top);
md_data <- mice(df_raw_top,m=1,maxit=5,meth='pmm',seed=500);
df_raw_top <- complete(md_data,1);

df_top <- df_raw_top
save(df_top,file="df_top.Rda")
#df_top <- load(file="df_top.Rda")

##########################
### Load Bottom Cohort ###
##########################
print("Reading and formatting data frames...")
  
# Read tab-delimited input using header row labels and interpret Python None strings as R NA values
df_raw_bottom = tbl_df(read.csv("/Users/jwang/Desktop/Results/2010-2013_unmatched_patient_feature_matrix_bottom_performing.csv",header=TRUE,sep=",",na.strings="None", stringsAsFactors=FALSE));

# Convert T/F = 1/0 (Columns labeled under binary_columns)
for (col_id in c(22:61)) {
  df_raw_bottom[,col_id] <- factor(ifelse(df_raw_bottom[,col_id]=="True", 1, 0))
}
# Condense dataframe
columns_of_interest = c(1, 6:296) # id, relevant covariates, label columns
df_raw_bottom <- df_raw_bottom[,columns_of_interest]

# Ensure legitimate variable names
colnames(df_raw_bottom) <- make.names(colnames(df_raw_bottom))

###################
### Missingness ###
###################

# Remove features with too much NA/missing data (by manual inspection)
df_raw_bottom = df_raw_bottom[,c(1:3,6:292)]

# Assess lab tests for percentage na: columns 7-15 in df
for (i in 7:15) {
  ratio_na = sum(is.na(df_raw_bottom[,i]))/dim(df_raw_bottom)[1]
  print(colnames(df_raw_bottom)[i])
  print(ratio_na)
}

# Remove urine, column 7 (missingness > 20%)
df_raw_bottom = df_raw_bottom[,c(1:6, 8:290)]

# Remove TT and admission diagnosis columns that are all zero
df_raw_bottom <- data.frame(df_raw_bottom)
# zero_cols = c()
# for (i in c(43:54, 57:284)) { # colindex 55 = income, colindex 56 = age
  # replace NA with 0
#   print(i)
#   df_raw_bottom[,i][is.na(df_raw_bottom)] <- 0
#   if ((length(df_raw_bottom[,i][df_raw_bottom[,i] == 1])) == 0) {
#     print(i)
#     print(colnames(df_raw_bottom)[i])
#     zero_cols <- c(zero_cols,i)
#   }
# }
df_raw_bottom <- df_raw_bottom[,-zero_cols]
df_raw_bottom$Charlson.Dementia.pre[is.na(df_raw_bottom$Charlson.Dementia.pre)] <- 0

# Convert remaining admission diagnosis columns to factor
for (i in c(56:185)) {
  df_raw_bottom[,i] = as.factor(df_raw_bottom[,i])
}

# Impute the remaining NA values: predictive mean imputation
md.pattern(df_raw_bottom);
md_data <- mice(df_raw_bottom,m=1,maxit=5,meth='pmm',seed=500);
df_raw_bottom <- complete(md_data,1);

df_bottom <- df_raw_bottom
save(df_bottom,file="df_bottom.Rda")
#df_bottom <- load(file="df_bottom.Rda")
```

PART 2: PROPENSITY SCORING AND MATCHING (TOP VS. BOTTOM)
Note: this module only considers the case if 2-way matching (1:1)
```{r}
##########################
### Propensity Scoring ###
##########################

# Add additional label column (we will be predicting probability of being in the top patient cohort)
df <- rbind(cbind(df_top[,c(1:55,208:212)],label=1),cbind(df_bottom[,c(1:55,208:212)],label=0))

print("Building propensity score model...")
# Logistic Regression
# Based on propensity of being assigned to the top patient cohort
id_column <- 1
label_columns <- c(56:61)
label <- as.factor(df$label)
glm.fit <- glm(label ~.,family=binomial(link='logit'), data=df[,c(-id_column, -label_columns)])

# Propensity Estimation
top_features <- df %>%
  filter(label == 1)
top_features <- top_features[,c(-id_column, -label_columns)] # remove id, label

bottom_features <- df %>%
  filter(label == 0)
bottom_features <- bottom_features[,c(-id_column, -label_columns)] # remove id, label

pred_top <- predict(glm.fit, newdata=top_features, type="response")
pred_bottom <- predict(glm.fit, newdata=bottom_features, type="response")
preds <- c(pred_top, pred_bottom)

####################
### Overlap Plot ###
####################
print("Plotting overlap...")
# note that logit = log(p/(1-p))
require(car)
logit_top <- as.numeric(logit(pred_top))
logit_bottom <- as.numeric(logit(pred_bottom))
histdata <- data.frame(rbind.data.frame(cbind.data.frame(logit=logit_top,label=1), cbind.data.frame(logit=logit_bottom,label=0)))
plot <- ggplot(histdata, aes(x=logit, fill=label)) + geom_histogram(alpha=0.2, position="identity")
plot

########################
### Caliper Matching ###
########################
print("Conducting caliper matching...")
# caliper = the maximum tolerated difference between matched subjects in a "non-perfect" matching 
# conduct matching w/o replacement on logit of propensity
temp <- data.frame(rbind.data.frame(cbind.data.frame(logit=logit_top,label=1), cbind.data.frame(logit=logit_bottom,label=0)))

# compute caliper: 0.2*sd(logit(propensity_scores))
propensity_scores <- temp$logit
opt_caliper <- 0.2*sd(propensity_scores)
print(opt_caliper)

#matches <- Match(X=temp$logit, Tr=temp$label, caliper=0.25, replace=FALSE, M=2)
matches <- Match(X=temp$logit, Tr=temp$label, caliper=opt_caliper, replace=F, M=1)

# number of patients in matched set
treated <- unique(matches$index.treated) # Top
control <- unique(matches$index.control) # Bottom
print("Post-Matching Cohort Sizes:")
length(treated)
length(control)

# save matched features in preparation for matching against the everyone cohort
top_matched_features <- df[matches$index.treated, ]
top_matched_features <- top_matched_features[,c(-id_column, -label_columns)] # remove id, label

bottom_matched_features <- df[matches$index.control, ]
bottom_matched_features <- bottom_matched_features[,c(-id_column, -label_columns)] # remove id, label

df_top_matched <- df_top[matches$index.treated,]

##############################
### Post-Matching Analysis ###
##############################

print("Preparing post-matching results...")

# Examine Matching Results
top_full <- df %>% 
  filter(label == 1)

bottom_full <- df %>% 
  filter(label == 0)

df_full <- rbind.data.frame(top_full, bottom_full)
# treated = top (1), control = bottom (0)
df_matched <- rbind.data.frame(df_full[treated,], df_full[control,])

# Output to CSV
write.csv(df_matched, file = "/Users/jwang/Desktop/Results/matched_top_bottom.csv")

# Before and After Comparisons
top_matched <- df_matched %>%
  filter(label == 1)
top_matched <- top_matched[,c(-id_column, -label_columns)]
bottom_matched <- df_matched %>%
  filter(label == 0)
bottom_matched <- bottom_matched[,c(-id_column, -label_columns)]

top_unmatched <-top_features
bottom_unmatched <- bottom_features

# Pre and Post Matching Standardized Difference in Means
print("Plotting pre- vs. post-matching SMD values...")

t1 <- cbind(top_unmatched, label=1)
t2 <- cbind(bottom_unmatched, label=0)
unmatched <- rbind(t1, t2)
unmatched_tab <- CreateTableOne(vars = colnames(unmatched), strata = "label", data = unmatched, test = FALSE)
print(unmatched_tab, smd = TRUE)

t1 <- cbind(top_matched, label=1)
t2 <- cbind(bottom_matched, label=0)
matched <- rbind(t1, t2)
matched_tab <- CreateTableOne(vars = colnames(matched), strata = "label", data = matched, test = FALSE)
print(matched_tab, smd = TRUE)

## Construct a data frame containing variable name and SMD from all methods
data_plot <- data.frame(variable  = names(ExtractSmd(unmatched_tab)),
                       Unmatched = ExtractSmd(unmatched_tab),
                       Matched   = ExtractSmd(matched_tab))

## Create long-format data for ggplot2
data_plot_melt <- melt(data = data_plot,
                     id.vars = c("variable"),
                     variable.name = "Method",
                     value.name = "SMD")

## Order variable names by magnitude of SMD
varNames <- as.character(data_plot$variable)[order(data_plot$Unmatched)]

## Order factor levels in the same order
data_plot_melt$variable <- factor(data_plot_melt$variable,
                                levels = varNames)

## Plot using ggplot2
# If including all covariates possible: use this data_plot_melt[-c(55:207,262:414),]
plot <- ggplot(data = data_plot_melt[-c(55,110),], mapping = aes(x = variable, y = SMD,
                                          group = Method, color = Method)) +
geom_line() +
geom_point() +
geom_hline(yintercept = 0.1, color = "black", size = 0.1) +
coord_flip() +
theme_bw() + theme(legend.key = element_blank())
print(plot)

# Pre and Post-Matching P-Values
print("Printing pre- vs. post-matching p-values...")

for (i in 1:length(top_features)) {
  print(colnames(top_matched)[i])
  
  tryCatch(if (length(unique(top_matched[,i])) > 2) { #mean
    print("Unmatched:")
    print("Top")
    print(mean(top_unmatched[,i]))
    #print(sd(top_unmatched[,i]))
    
    print("Bottom")
    print(mean(bottom_unmatched[,i]))
    #print(sd(bottom_unmatched[,i]))
    #print(wilcox.test(expert_unmatched[,i], everyone_unmatched[,i], alternative = "two.sided")$p.value)
    print("P-value:")
    print(t.test(top_unmatched[,i], bottom_unmatched[,i])$p.value)
    
    print("Matched:")
    print("Top")
    print(mean(top_matched[,i]))
    #print(sd(top_matched[,i]))
    
    print("Bottom")
    print(mean(bottom_matched[,i]))
    #print(sd(bottom_matched[,i]))
    #print(wilcox.test(expert_matched[,i], everyone_matched[,i], alternative = "two.sided")$p.value)
    
    print("P-value:")
    print(t.test(top_matched[,i], bottom_matched[,i])$p.value)
  }
  else { # fraction    
    print("Unmatched:")
    print("Top")
    print(length(top_unmatched[,i][top_unmatched[,i] == 1])/dim(top_unmatched)[1])
    
    print("Bottom")
    print(length(bottom_unmatched[,i][bottom_unmatched[,i] == 1])/dim(bottom_unmatched)[1])
    #print(chisq.test(table(expert_unmatched[,i], everyone_unmatched[,i]))$p.value)
    
    print("P-value:")
    print(t.test(as.numeric(top_unmatched[,i]), as.numeric(bottom_unmatched[,i]))$p.value)

    print("Matched:")
    print("Top")
    print(length(top_matched[,i][top_matched[,i] == 1])/dim(top_matched)[1])
    
    print("Bottom")
    print(length(bottom_matched[,i][bottom_matched[,i] == 1])/dim(bottom_matched)[1])
    #print(chisq.test(table(expert_matched[,i], everyone_matched[,i]))$p.value)
    
    print("P-value:")
    print(t.test(as.numeric(top_matched[,i]), as.numeric(bottom_matched[,i]))$p.value)
  }, error=function(e){})
}
```
PART 3: LOAD IN ALL PATIENTS ("EVERYONE") FROM 2010-2013
```{r}
##########################
### Load Bottom Cohort ###
##########################
print("Reading and formatting data frames...")
  
# Read tab-delimited input using header row labels and interpret Python None strings as R NA values
df_raw_everyone = tbl_df(read.csv("/Users/jwang/Desktop/Results/2010-2013_patient_feature_matrix.csv",header=TRUE,sep=",",na.strings="None", stringsAsFactors=FALSE));

# Convert T/F = 1/0 (Columns labeled under binary_columns)
for (col_id in c(22:61)) {
  df_raw_everyone[,col_id] <- factor(ifelse(df_raw_everyone[,col_id]=="True", 1, 0))
}
# Condense dataframe
columns_of_interest = c(1, 6:296) # id, relevant covariates, label columns
df_raw_everyone <- df_raw_everyone[,columns_of_interest]

# Ensure legitimate variable names
colnames(df_raw_everyone) <- make.names(colnames(df_raw_everyone))

###################
### Missingness ###
###################

# Remove features with too much NA/missing data (by manual inspection)
df_raw_everyone = df_raw_everyone[,c(1:3,6:292)]

# Assess lab tests for percentage na: columns 7-15 in df
for (i in 7:15) {
  ratio_na = sum(is.na(df_raw_everyone[,i]))/dim(df_raw_everyone)[1]
  print(colnames(df_raw_everyone)[i])
  print(ratio_na)
}

# Remove urine, column 7 (missingness > 20%)
df_raw_everyone = df_raw_everyone[,c(1:6, 8:290)]

# Remove TT and admission diagnosis columns that are all zero
df_raw_everyone <- data.frame(df_raw_everyone)
# zero_cols = c()
# for (i in c(43:54, 57:284)) { # colindex 55 = income, colindex 56 = age
  # replace NA with 0
#   print(i)
#   df_raw_bottom[,i][is.na(df_raw_bottom)] <- 0
#   if ((length(df_raw_bottom[,i][df_raw_bottom[,i] == 1])) == 0) {
#     print(i)
#     print(colnames(df_raw_bottom)[i])
#     zero_cols <- c(zero_cols,i)
#   }
# }
df_raw_everyone <- df_raw_everyone[,-zero_cols]
df_raw_everyone$Charlson.Dementia.pre[is.na(df_raw_everyone$Charlson.Dementia.pre)] <- 0

# Convert remaining admission diagnosis columns to factor
for (i in c(56:185)) {
  df_raw_everyone[,i] = as.factor(df_raw_everyone[,i])
}

# Impute the remaining NA values: predictive mean imputation
md.pattern(df_raw_everyone);
md_data <- mice(df_raw_everyone,m=1,maxit=5,meth='pmm',seed=500);
df_raw_everyone <- complete(md_data,1);

df_everyone <- df_raw_everyone
save(df_everyone,file="df_everyone.Rda")
#df_everyone <- load(file="df_everyone.Rda")
```

PART 4: PROPENSITY SCORING AND MATCHING (TOP* VS. EVERYONE)
TOP* = already matched top patient cohort from PART 3
```{r}
##########################
### Propensity Scoring ###
##########################

# Instead of training the model on all >60,000 patients which adds a lot of noise, reduce the training set by 10-fold
df <- rbind(cbind(df_top_matched[,c(1:55,208:212)],label=1),cbind(df_everyone[1:10000,c(1:55,208:212)],label=0))

print("Building propensity score model...")
# Logistic Regression
id_column <- 1
#label_columns <- c(208:213)
label_columns <- c(56:61)
label <- as.factor(df$label)
glm.fit <- glm(label ~.,family=binomial(link='logit'), data=df[,c(-id_column, -label_columns)])

# Propensity Estimation
top_matched_features <- df %>%
  filter(label == 1)
top_matched_features <- top_matched_features[,c(-id_column, -label_columns)] # remove id, label

everyone_features <- df %>%
  filter(label == 0)
everyone_features <- everyone_features[,c(-id_column, -label_columns)] # remove id, label

pred_top_matched <- predict(glm.fit, newdata=top_matched_features, type="response")
pred_everyone <- predict(glm.fit, newdata=everyone_features, type="response")
preds <- c(pred_top_matched, pred_everyone)

####################
### Overlap Plot ###
####################
print("Plotting overlap...")
# note that logit = log(p/(1-p))
require(car)
logit_top_matched <- as.numeric(logit(pred_top_matched))
logit_everyone <- as.numeric(logit(pred_everyone))
histdata <- data.frame(rbind.data.frame(cbind.data.frame(logit=logit_top_matched,label=1), cbind.data.frame(logit=logit_everyone,label=0)))
plot <- ggplot(histdata, aes(x=logit, fill=label)) + geom_histogram(alpha=0.2, position="identity")
plot

########################
### Caliper Matching ###
########################
print("Conducting caliper matching...")
# caliper = the maximum tolerated difference between matched subjects in a "non-perfect" matching 
# conduct matching w/o replacement on logit of propensity
temp <- data.frame(rbind.data.frame(cbind.data.frame(logit=logit_top_matched,label=1), cbind.data.frame(logit=logit_everyone,label=0)))

# compute caliper: 0.2*sd(logit(propensity_scores))
propensity_scores <- temp$logit
opt_caliper <- 0.2*sd(propensity_scores)
print(opt_caliper)

#matches <- Match(X=temp$logit, Tr=temp$label, caliper=0.25, replace=FALSE, M=2)
matches <- Match(X=temp$logit, Tr=temp$label, caliper=opt_caliper, replace=F, M=5)

# number of patients in matched set
treated <- unique(matches$index.treated) # Top Matched
control <- unique(matches$index.control) # Everyone
print("Post-Matching Cohort Sizes:")
length(treated)
length(control)

##############################
### Post-Matching Analysis ###
##############################

print("Preparing post-matching results...")

# Examine Matching Results
top_matched_full <- df %>% 
  filter(label == 1)

everyone_full <- df %>% 
  filter(label == 0)

df_full <- rbind.data.frame(top_matched_full, everyone_full)
# treated = top (1), control = everyone (0)
df_matched <- rbind.data.frame(df_full[treated,], df_full[control,])

# Output to CSV
write.csv(df_matched, file = "/Users/jwang/Desktop/Results/matched_top_everyone.csv")

# Before and After Comparisons
top_matched_matched <- df_matched %>%
  filter(label == 1)
top_matched_matched <- top_matched_matched[,c(-id_column, -label_columns)]
everyone_matched <- df_matched %>%
  filter(label == 0)
everyone_matched <- everyone_matched[,c(-id_column, -label_columns)]

top_matched_unmatched <-top_matched_features
everyone_unmatched <- everyone_features

# Pre and Post Matching Standardized Difference in Means
print("Plotting pre- vs. post-matching SMD values...")

t1 <- cbind(top_matched_unmatched, label=1)
t2 <- cbind(everyone_unmatched, label=0)
unmatched <- rbind(t1, t2)
unmatched_tab <- CreateTableOne(vars = colnames(unmatched), strata = "label", data = unmatched, test = FALSE)
print(unmatched_tab, smd = TRUE)

t1 <- cbind(top_matched_matched, label=1)
t2 <- cbind(everyone_matched, label=0)
matched <- rbind(t1, t2)
matched_tab <- CreateTableOne(vars = colnames(matched), strata = "label", data = matched, test = FALSE)
print(matched_tab, smd = TRUE)

## Construct a data frame containing variable name and SMD from all methods
data_plot <- data.frame(variable  = names(ExtractSmd(unmatched_tab)),
                       Unmatched = ExtractSmd(unmatched_tab),
                       Matched   = ExtractSmd(matched_tab))

## Create long-format data for ggplot2
data_plot_melt <- melt(data = data_plot,
                     id.vars = c("variable"),
                     variable.name = "Method",
                     value.name = "SMD")

## Order variable names by magnitude of SMD
varNames <- as.character(data_plot$variable)[order(data_plot$Unmatched)]

## Order factor levels in the same order
data_plot_melt$variable <- factor(data_plot_melt$variable,
                                levels = varNames)

## Plot using ggplot2
plot <- ggplot(data = data_plot_melt[-c(55,110),], mapping = aes(x = variable, y = SMD,
                                          group = Method, color = Method)) +
geom_line() +
geom_point() +
geom_hline(yintercept = 0.1, color = "black", size = 0.1) +
coord_flip() +
theme_bw() + theme(legend.key = element_blank())
print(plot)

# Pre and Post-Matching P-Values
print("Printing pre- vs. post-matching p-values...")

for (i in 1:length(top_features)) {
  print(colnames(top_matched)[i])
  
  tryCatch(if (length(unique(top_matched[,i])) > 2) { #mean
    print("Unmatched:")
    print("Top*")
    print(mean(top_matched_unmatched[,i]))
    #print(sd(top_unmatched[,i]))
    
    print("Everyone")
    print(mean(everyone_unmatched[,i]))
    #print(sd(bottom_unmatched[,i]))
    #print(wilcox.test(expert_unmatched[,i], everyone_unmatched[,i], alternative = "two.sided")$p.value)
    print("P-value:")
    print(t.test(top_matched_unmatched[,i], everyone_unmatched[,i])$p.value)
    
    print("Matched:")
    print("Top")
    print(mean(top_matched_matched[,i]))
    #print(sd(top_matched[,i]))
    
    print("Everyone")
    print(mean(everyone_matched[,i]))
    #print(sd(bottom_matched[,i]))
    #print(wilcox.test(expert_matched[,i], everyone_matched[,i], alternative = "two.sided")$p.value)
    
    print("P-value:")
    print(t.test(top_matched_matched[,i], everyone_matched[,i])$p.value)
  }
  else { # fraction    
    print("Unmatched:")
    print("Top")
    print(length(top_matched_unmatched[,i][top_matched_unmatched[,i] == 1])/dim(top_matched_unmatched)[1])
    
    print("Everyone")
    print(length(everyone_unmatched[,i][everyone_unmatched[,i] == 1])/dim(everyone_unmatched)[1])
    #print(chisq.test(table(expert_unmatched[,i], everyone_unmatched[,i]))$p.value)
    
    print("P-value:")
    print(t.test(as.numeric(top_matched_unmatched[,i]), as.numeric(everyone_unmatched[,i]))$p.value)

    print("Matched:")
    print("Top")
    print(length(top_matched_matched[,i][top_matched_matched[,i] == 1])/dim(top_matched_matched)[1])
    
    print("Everyone")
    print(length(everyone_matched[,i][everyone_matched[,i] == 1])/dim(everyone_matched)[1])
    #print(chisq.test(table(expert_matched[,i], everyone_matched[,i]))$p.value)
    
    print("P-value:")
    print(t.test(as.numeric(top_matched_matched[,i]), as.numeric(everyone_matched[,i]))$p.value)
  }, error=function(e){})
}
```
PART 5: COMPARING EVERYONE AND BOTTOM
```{r}
##############################
### Post-Matching Analysis ###
##############################

print("Preparing post-matching results...")

# Before and After Comparisons

#bottom_matched and everyone_matched have been created in previous parts
bottom_unmatched <- bottom_features
everyone_unmatched <- everyone_features

# Pre and Post Matching Standardized Difference in Means
print("Plotting pre- vs. post-matching SMD values...")

t1 <- cbind(bottom_unmatched, label=1)
t2 <- cbind(everyone_unmatched, label=0)
unmatched <- rbind(t1, t2)
unmatched_tab <- CreateTableOne(vars = colnames(unmatched), strata = "label", data = unmatched, test = FALSE)
print(unmatched_tab, smd = TRUE)

t1 <- cbind(bottom_matched, label=1)
t2 <- cbind(everyone_matched, label=0)
matched <- rbind(t1, t2)
matched_tab <- CreateTableOne(vars = colnames(matched), strata = "label", data = matched, test = FALSE)
print(matched_tab, smd = TRUE)

## Construct a data frame containing variable name and SMD from all methods
data_plot <- data.frame(variable  = names(ExtractSmd(unmatched_tab)),
                       Unmatched = ExtractSmd(unmatched_tab),
                       Matched   = ExtractSmd(matched_tab))

## Create long-format data for ggplot2
data_plot_melt <- melt(data = data_plot,
                     id.vars = c("variable"),
                     variable.name = "Method",
                     value.name = "SMD")

## Order variable names by magnitude of SMD
varNames <- as.character(data_plot$variable)[order(data_plot$Unmatched)]

## Order factor levels in the same order
data_plot_melt$variable <- factor(data_plot_melt$variable,
                                levels = varNames)

## Plot using ggplot2
plot <- ggplot(data = data_plot_melt[-c(55,110),], mapping = aes(x = variable, y = SMD,
                                          group = Method, color = Method)) +
geom_line() +
geom_point() +
geom_hline(yintercept = 0.1, color = "black", size = 0.1) +
coord_flip() +
theme_bw() + theme(legend.key = element_blank())
print(plot)

# Pre and Post-Matching P-Values
print("Printing pre- vs. post-matching p-values...")

for (i in 1:length(bottom_features)) {
  print(colnames(bottom_matched)[i])
  
  tryCatch(if (length(unique(bottom_matched[,i])) > 2) { #mean
    print("Unmatched:")
    print("Bottom")
    print(mean(bottom_unmatched[,i]))
    #print(sd(top_unmatched[,i]))
    
    print("Everyone")
    print(mean(everyone_unmatched[,i]))
    #print(sd(bottom_unmatched[,i]))
    #print(wilcox.test(expert_unmatched[,i], everyone_unmatched[,i], alternative = "two.sided")$p.value)
    print("P-value:")
    print(t.test(bottom_unmatched[,i], everyone_unmatched[,i])$p.value)
    
    print("Matched:")
    print("Bottom")
    print(mean(bottom_matched[,i]))
    #print(sd(top_matched[,i]))
    
    print("Everyone")
    print(mean(everyone_matched[,i]))
    #print(sd(bottom_matched[,i]))
    #print(wilcox.test(expert_matched[,i], everyone_matched[,i], alternative = "two.sided")$p.value)
    
    print("P-value:")
    print(t.test(bottom_matched[,i], everyone_matched[,i])$p.value)
  }
  else { # fraction    
    print("Unmatched:")
    print("Bottom")
    print(length(bottom_unmatched[,i][bottom_unmatched[,i] == 1])/dim(bottom_unmatched)[1])
    
    print("Everyone")
    print(length(everyone_unmatched[,i][everyone_unmatched[,i] == 1])/dim(everyone_unmatched)[1])
    #print(chisq.test(table(expert_unmatched[,i], everyone_unmatched[,i]))$p.value)
    
    print("P-value:")
    print(t.test(as.numeric(bottom_unmatched[,i]), as.numeric(everyone_unmatched[,i]))$p.value)

    print("Matched:")
    print("Bottom")
    print(length(bottom_matched[,i][bottom_matched[,i] == 1])/dim(bottom_matched)[1])
    
    print("Everyone")
    print(length(everyone_matched[,i][everyone_matched[,i] == 1])/dim(everyone_matched)[1])
    #print(chisq.test(table(expert_matched[,i], everyone_matched[,i]))$p.value)
    
    print("P-value:")
    print(t.test(as.numeric(bottom_matched[,i]), as.numeric(everyone_matched[,i]))$p.value)
  }, error=function(e){})
}
```